---
title: "Survival Model analysis"
output: html_document
date: "2024-11-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}

# Load necessary libraries

library(knitr)
library(kableExtra)

library(survival)
library(MASS)
# Install necessary packages if not already installed
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")
if (!requireNamespace("tidyr", quietly = TRUE)) install.packages("tidyr")

# Load packages
library(dplyr)
library(tidyr)

if (!requireNamespace("nortest", quietly = TRUE)) install.packages("nortest")
library(nortest)

# Load necessary libraries
library(survival)
library(survminer)

```


```{r}
# Set working directory to where 'cleaned_final_cox_dataset.csv' is located
setwd("D:/LLY/Capstone")

# Load dataset
data <- read.csv("cleaned_final_cox_dataset.csv")
```


```{r}
# Ensure the dataset's structure is correct for a Cox model
str(data)
```

```{r}
# Convert character-based categorical variables to factors
data$insurance_Medicare <- as.factor(data$insurance_Medicare)
data$admission_location_EMERGENCY.ROOM.ADMIT <- as.factor(data$admission_location_EMERGENCY.ROOM.ADMIT)
data$admission_location_PHYS.REFERRAL.NORMAL.DELI <- as.factor(data$admission_location_PHYS.REFERRAL.NORMAL.DELI)
data$admission_location_CLINIC.REFERRAL.PREMATURE <- as.factor(data$admission_location_CLINIC.REFERRAL.PREMATURE)
data$admission_location_TRANSFER.FROM.HOSP.EXTRAM <- as.factor(data$admission_location_TRANSFER.FROM.HOSP.EXTRAM)
data$admission_type_URGENT <- as.factor(data$admission_type_URGENT)
data$is_weekend_admission <- as.factor(data$is_weekend_admission)

```


```{r}


# List of categorical variables
categorical_vars <- c(
  "insurance_Medicare", 
  "admission_location_EMERGENCY.ROOM.ADMIT",
  "admission_location_PHYS.REFERRAL.NORMAL.DELI",
  "admission_location_CLINIC.REFERRAL.PREMATURE",
  "admission_location_TRANSFER.FROM.HOSP.EXTRAM",
  "admission_type_URGENT",
  "is_weekend_admission"
)

# Kaplan-Meier survival curves for each categorical variable
for (variable_name in categorical_vars) {
  # Ensure the variable is a factor
  data[[variable_name]] <- as.factor(data[[variable_name]])
  
  # Fit a Kaplan-Meier survival curve
  km_fit <- survfit(as.formula(paste("Surv(los_x, expire_flag) ~", variable_name)), data = data)
  
  # Plot the survival curve without risk table
  plot <- ggsurvplot(
    km_fit,
    data = data,
    risk.table = FALSE,            # Disable risk table
    pval = TRUE,                   # Show p-value from log-rank test
    conf.int = TRUE,               # Show confidence intervals
    legend.title = variable_name,  # Use variable name as legend title
    legend.labs = levels(data[[variable_name]]), # Use levels of the variable as legend labels
    xlab = "Time (days)",          # Label for x-axis
    ylab = "Survival Probability", # Label for y-axis
    title = paste("Kaplan-Meier Survival Curve by", variable_name)
  )
  
  # Print the plot to ensure rendering
  print(plot)
}



```
The Kaplan-Meier survival curves indicate statistically significant differences in survival probabilities across various groups (p < 0.05 for all comparisons). Key observations include: patients with Medicare insurance demonstrate higher survival probabilities, suggesting a protective effect potentially linked to better access to care. Admission through the emergency room and physician referrals are associated with improved survival, likely due to faster or more structured medical interventions. Patients transferred from other hospitals also show higher survival rates, reflecting potentially better resource allocation. While urgent admissions exhibit a smaller survival advantage, the difference remains statistically significant. These findings highlight the importance of insurance status, admission methods, and referral pathways in influencing patient survival, warranting further validation through multivariate Cox regression analysis.


```{r}


# Get all column names from the dataset
all_columns <- colnames(data)

# Initialize a list to store Cox regression results
cox_results <- list()

# Perform univariate Cox regression on all variables
for (var in all_columns) {
  # Skip time variable and target variable
  if (var %in% c("los_x", "expire_flag")) next
  
  # Create the Cox regression formula
  formula <- as.formula(paste("Surv(los_x, expire_flag) ~", var))
  
  # Attempt to run the Cox regression
  tryCatch({
    cox_model <- coxph(formula, data = data)
    cox_summary <- summary(cox_model)
    
    # Extract results and store them
    cox_results[[var]] <- data.frame(
      Variable = var,
      HR = exp(cox_summary$coefficients[, "coef"]),  # Hazard ratio
      CI_lower = exp(cox_summary$conf.int[, "lower .95"]),  # Lower confidence interval (95%)
      CI_upper = exp(cox_summary$conf.int[, "upper .95"]),  # Upper confidence interval (95%)
      P_value = cox_summary$coefficients[, "Pr(>|z|)"]  # P-value
    )
  }, error = function(e) {
    message(paste("Error processing", var, ":", e$message))
  })
}

# Combine all results into a single data frame
cox_results_df <- do.call(rbind, cox_results)

# Print all Cox regression results
print("Cox Regression Results:")
print(cox_results_df)

# Filter significant variables (P-value < 0.05)
significant_vars <- cox_results_df[cox_results_df$P_value < 0.05, ]

# Print significant variables
print("Significant Variables (P < 0.05):")
print(significant_vars)



```
```{r}
unadjusted_model <- coxph(Surv(los_x, expire_flag) ~ is_weekend_admission , data = data)
summary(unadjusted_model)
```


```{r}


# Re-fit the Cox model, forcing the inclusion of is_weekend_admission
cox_model <- coxph(Surv(los_x, expire_flag) ~ is_weekend_admission + ., data = data)

# Run stepwise selection with is_weekend_admission forced to remain
library(MASS)
stepwise_model <- stepAIC(cox_model, scope = list(lower = ~ is_weekend_admission, upper = ~ is_weekend_admission + .), direction = "both", trace = 0)

# Summarize the final model
summary(stepwise_model)

```

```{r}
# Get the coefficient summary of the stepwise_model
stepwise_summary <- summary(stepwise_model)

# Extract p-values
stepwise_p_values <- stepwise_summary$coefficients[, "Pr(>|z|)"]

# Filter significant variables (p < 0.05)
significant_vars <- names(stepwise_p_values[stepwise_p_values < 0.05])

# Print significant variables
print("Significant variables from stepwise_model:")
print(significant_vars)

```
```{r}
# Remove "True" from variable names
significant_vars_cleaned <- gsub("True", "", significant_vars)

# Ensure the inclusion of 'is_weekend_admission' variable
significant_vars_cleaned <- union(significant_vars_cleaned, "is_weekend_admission")

# Construct a new formula
significant_formula <- as.formula(
  paste("Surv(los_x, expire_flag) ~", paste(significant_vars_cleaned, collapse = " + "))
)

# Fit a Cox model using the cleaned variables
reduced_model <- coxph(significant_formula, data = data)

# Output the summary of the Reduced Model
summary(reduced_model)


```

```{r}
# Assign the stepwise_model as the full model
full_model <- stepwise_model

# Compare the reduced model with the full model using Likelihood Ratio Test (LRT)
lrt_full_vs_reduced <- anova(full_model, reduced_model, test = "LRT")

# Compare the unadjusted model with the full model using LRT
lrt_unadjusted_vs_full <- anova(unadjusted_model, full_model, test = "LRT")

# Compare the unadjusted model with the reduced model using LRT
lrt_unadjusted_vs_reduced <- anova(unadjusted_model, reduced_model, test = "LRT")

# Print the results
print("Likelihood Ratio Test (Full Model vs Reduced Model):")
print(lrt_full_vs_reduced)

print("Likelihood Ratio Test (Unadjusted Model vs Full Model):")
print(lrt_unadjusted_vs_full)

print("Likelihood Ratio Test (Unadjusted Model vs Reduced Model):")
print(lrt_unadjusted_vs_reduced)



```
```{r}
# Compare Full Model vs Reduced Model using ANOVA
anova_full_vs_reduced <- anova(full_model, reduced_model, test = "LRT")

# Compare Full Model vs Unadjusted Model using ANOVA
anova_full_vs_unadjusted <- anova(full_model, unadjusted_model, test = "LRT")

# Compare Reduced Model vs Unadjusted Model using ANOVA
anova_reduced_vs_unadjusted <- anova(reduced_model, unadjusted_model, test = "LRT")

# Print the results
print("ANOVA: Full Model vs Reduced Model")
print(anova_full_vs_reduced)

print("ANOVA: Full Model vs Unadjusted Model")
print(anova_full_vs_unadjusted)

print("ANOVA: Reduced Model vs Unadjusted Model")
print(anova_reduced_vs_unadjusted)

```
```{r}
# Calculate AIC and BIC for each model

# AIC and BIC for Full Model
aic_full <- AIC(full_model)
bic_full <- BIC(full_model)

# AIC and BIC for Reduced Model
aic_reduced <- AIC(reduced_model)
bic_reduced <- BIC(reduced_model)

# AIC and BIC for Unadjusted Model
aic_unadjusted <- AIC(unadjusted_model)
bic_unadjusted <- BIC(unadjusted_model)

# Print the results
print("AIC and BIC for Full Model:")
print(paste("AIC:", aic_full))
print(paste("BIC:", bic_full))

print("AIC and BIC for Reduced Model:")
print(paste("AIC:", aic_reduced))
print(paste("BIC:", bic_reduced))

print("AIC and BIC for Unadjusted Model:")
print(paste("AIC:", aic_unadjusted))
print(paste("BIC:", bic_unadjusted))

```

Model1: Surv(los_x, expire_flag) ~ is_weekend_admission

Model2: Surv(los_x, expire_flag) ~ is_weekend_admission + positiveculture + gcs + gcsmotor + lactate + bloodureanitrogen + hemoglobin + intnormalisedratio + chloride + hematocrit + age_years + insurance_Medicare + gcsverbal + admission_location_EMERGENCY.ROOM.ADMIT + admission_location_PHYS.REFERRAL.NORMAL.DELI + admission_location_CLINIC.REFERRAL.PREMATURE + admission_location_TRANSFER.FROM.HOSP.EXTRAM

Model3: Surv(los_x, expire_flag) ~ positiveculture + gcs + gcsmotor + lactate + bloodureanitrogen + intnormalisedratio + chloride + hematocrit + age_years + insurance_Medicare + gcsverbal + admission_location_PHYS.REFERRAL.NORMAL.DELI + admission_location_CLINIC.REFERRAL.PREMATURE + admission_location_TRANSFER.FROM.HOSP.EXTRAM


```{r}

# 创建模型对比数据框
model_comparison <- data.frame(
  Term = c(
    "Intercept",
    "is_weekend_admission",
    "positiveculture",
    "gcs",
    "gcsmotor",
    "lactate",
    "bloodureanitrogen",
    "hemoglobin",
    "intnormalisedratio",
    "chloride",
    "hematocrit",
    "age_years",
    "insurance_Medicare",
    "AIC",
    "BIC"
  ),
  Model1 = c(
    "0.9858*** (0.938, 1.036)", "-", "-", "-", "-", "-", "-", "-", "-", "-", "-", "-", "-", 
    "140726", "140733"
  ),
  Model2 = c(
    "0.9691 (0.9219, 1.0188)", "0.9691*** (0.9219, 1.0188)", "1.0583* (1.0029, 1.1168)", 
    "1.0452** (1.0113, 1.0803)", "0.9239*** (0.8804, 0.9696)", "1.0212* (1.0016, 1.0413)", 
    "1.0029*** (1.0018, 1.0040)", "-", "1.0191* (1.0020, 1.0365)", 
    "0.9887*** (0.9844, 0.9931)", "0.9827*** (0.9771, 0.9883)", "1.0231*** (1.0212, 1.0250)", 
    "1.1651*** (1.0958, 1.2387)", "138296", "138401"
  ),
  Model3 = c(
    "0.9690 (0.9218, 1.0186)", "0.9690*** (0.9218, 1.0186)", "1.0556* (1.0003, 1.1140)", 
    "1.0448** (1.0109, 1.0798)", "0.9247*** (0.8811, 0.9704)", "1.0211* (1.0015, 1.0411)", 
    "1.0028*** (1.0017, 1.0039)", "0.9792 (0.9512, 1.0081)", "1.0189* (1.0016, 1.0364)", 
    "0.9887*** (0.9843, 0.9931)", "0.9885* (0.9788, 0.9984)", "1.0230*** (1.0212, 1.0249)", 
    "1.1623*** (1.0932, 1.2358)", "138295", "138414"
  )
)

# 创建表格
kable(model_comparison, align = "lccc", booktabs = TRUE, caption = "Comparison of Three Cox Models (95% CI)") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  footnote(
    general = c(
      "*** p < 0.001 (Highly significant)",
      "** p < 0.01 (Moderately significant)",
      "* p < 0.05 (Significant at 5% level)",
      "(95% CI) refers to the 95% Confidence Interval, representing the range of plausible values for the parameter estimate."
    ),
    general_title = "Notes:",
    footnote_as_chunk = TRUE
  )


```

Model1 (Baseline model):
This is a univariate model, including only is_weekend_admission as the explanatory variable. It serves as a baseline to measure the impact of admission time (weekend vs. weekday) on survival time. Without adjusting for other covariates, it provides a limited understanding of the survival outcome.

Model2:
Building upon Model1, this model incorporates additional clinical and demographic variables, such as positiveculture, gcs, and age_years. The inclusion of these variables allows the model to capture more patient-specific differences, significantly improving its fit and explanatory power.

Model3:
This simplified multivariable model retains only the significant variables from Model2. By removing non-significant variables, it reduces complexity while maintaining high explanatory power. This approach enhances interpretability and avoids overfitting.


Univariate Effects (Model1):

Model1 considers the sole effect of weekend admissions (is_weekend_admission) on survival time.
The model’s poor fit indicates that a single variable is insufficient to comprehensively explain survival time variability.

Adjusted Effects (Model2):

Model2 demonstrates improved fit by including multiple clinical and demographic factors.
Key variables such as gcs, age_years, and bloodureanitrogen significantly influence survival time, highlighting their importance in outcome prediction.

Model Simplification (Model3):

Model3 reduces complexity by excluding non-significant variables from Model2.
This simplification improves interpretability and ensures that the model focuses on the most influential predictors.

Clinical and Statistical Implications:

Variables like age (age_years) and lab results (e.g., bloodureanitrogen and hematocrit) have significant impacts on survival outcomes after adjustment.
The effect of weekend admission (is_weekend_admission) becomes non-significant after adjusting for other covariates, suggesting its influence is either weak or explained by other variables.

From the Perspective of Model Fit
AIC and BIC are criteria for evaluating the goodness of fit of a model. Smaller values indicate better model fit. Among the three models:

Model3 has the lowest AIC and BIC, which suggests it provides the best fit to the data when more parameters are included.
However, when model complexity is considered, Model2 may be a more balanced choice.

Summary
Patient characteristics, such as age and laboratory markers, are critical in predicting survival outcomes.
Model simplification (Model3) achieves a balance between explanatory power and complexity, making it the preferred model for interpretation.
Fit statistics (AIC and BIC) indicate that Model3 has the best fit, validating its suitability as the optimal model for the data.
Significance testing identifies key predictors (e.g., gcs, bloodureanitrogen) that should be prioritized in clinical decision-making.
As model complexity increases, goodness of fit improves, but interpretability may decrease. Simplifying the model by removing non-significant variables is a practical trade-off.
These findings provide a robust statistical basis for understanding the data and guiding clinical practice.

```{r}
# Check the proportionality assumption
test_proportionality <- cox.zph(reduced_model)
print(test_proportionality)

```


```


